\chapter{Conclusion}

In this manuscript, I have addressed the following questions and answered them based on results to a suite of micro-benchmarks I created.  Those micro-benchmarks aimed at being representative of the current load faced by INGInious when dealing with student's submissions and bring to light how different configurations could affect the responsiveness of the platform.

\begin{enumerate}
  \item Compared to other available solutions, how good is the current configuration chosen by INGInious to face the responsiveness challenge of the platform?  How much better could it be?  How easy would it be to improve it?  Do some solutions involve tradeoffs in terms of maturity, support or maintainability?
  \item Could there be a solution tailor-made for the specific case of INGInious?  What would it be?  What would it cost to use it?
  \item What would be the cost of providing stronger/safer isolation to the containers used by INGInious?  Which opportunities could it bring?
\end{enumerate}

To respond to the first question, I have questioned the current configuration that INGInious has chosen to face the responsiveness challenge of the platform.  I have identified different changes that could be made to it and shown whether or not they would be worth it.  The most interesting change would be to change the current container runtime in use, \texttt{runc}, to \texttt{crun}, which performs as much as twice as fast in some tests.

For the second question, I have presented what would be the best solution for INGInious's case and measured its performance gain, in my micro-benchmarks, over the best existing alternative.  Showing that the difference we can still gain is significant, performing up to three times as fast as INGInious's current configuration.

And finally, for the last question, I have discussed some of the safer isolation solutions that are existing, virtualization and rootless containerization.  I presented new opportunities it creates for INGInious, like creating tasks where the student would have full root permissions inside of a container.  Here is a summary of part of the configuration I tried, and what they are worth in terms of performance, isolation, ease of use and support (through documentation, community, forums).

\begin{center}
  \begin{tabular}{|ccc|c|c|c|c|}
    \hline
    \textbf{Manager} & \textbf{Runtime} & \textbf{Rootless} & \textbf{Isol.}\footnotemark & \textbf{Resp.}\footnotemark & \textbf{Usab.}\footnotemark & \textbf{Support} \\
    \hline
    \hline
    Docker & runc & No & + & +++ & +++ & +++ \\
    Docker & crun & No & + & ++++ & +++ & +++\\
    Docker & kata-runtime\footnotemark & No & +++ & ++ & ++ & +++\\
    Docker & kata-fc\footnotemark & No & +++ & + & ++ & +++\\
    Podman & runc & No & + & ++ & +++ & +++\\
    Podman & crun & No & + & +++ & +++ & +++\\
    Podman & crun & Yes & ++ & ++ & +++ & +++\\
    LXD & LXC & No & + & +++ & +++ & ++\\
    \hline
  \end{tabular}
\end{center}
\footnotetext[1]{Isolation}
\footnotetext[2]{Responsiveness}
\footnotetext[3]{Usability, ease of use}
\footnotetext[4]{Kata Containers with Qemu hypervisor}
\footnotetext[5]{Kata Containers with Firecracker hypervisor}

As there is no bad solution, there is no bad score either, but the goal is to have as many "+" as possible.

\subsubsection{Personal enrichment}
A master thesis is much more than reading a bunch of papers, writing nice code and plotting some graphs.  It has been a real challenge to discover this whole new world that is containerization and I have learned many more things than what my results show.  It was my first time dealing with such a big project, and I was the only one directly contributing to it.  It required more planning than what I usually need.  It was also the first time that I was not relying on a course's support to learn new things. I had to find other trustful content and build my knowledge of the situation myself.  This project has also been a great opportunity for me to enjoy the greatness of open-source. All the tools I have handled are open-source, and I could get support from the community when I required it, which was a big help and made me feel less alone in this.

\subsubsection{Follow up questions}

Even a year of work is not enough for me to cover everything on the subject I presented here.  Partly because the containerization world is continuously growing and new things to consider appear regularly.  And partly because I started this year with no experience or even basic knowledge on the subject, I did not know from the start what I was going to do and going further in my work has not ceased to open me new doors to look behind.  Here is then a small enumeration of thing that, if I had one more year, I would consider going into:
\begin{enumerate}
  \item As the different tools I compared are continuously evolving, and new opportunities appear, the amount of configuration to consider grows really big.  It might be interesting to create a tool that could easily integrate into its benchmark new upcoming solutions.  It would allow us to get periodically updated results as some tools grow.
  \item LXD can also be used to manage virtual machines, I do not know how it works, what it relies on and if the performances are anywhere near ones of containers, but it could be worth taking a look at.
  \item LXD as also its own "rootless" containers feature, with unprivileged containers, I did not have the chance to take a look at this but it could be interesting as well.
  \item My knowledge of virtualization technologies is weaker than containerization one, and there is one result that I can not explain, which is the really bad behavior of storage drivers others than devicemapper with Kata Containers.
\end{enumerate}
